\chapter{Foreword} % Main chapter title

\label{Foreword} % For referencing the chapter elsewhere, use \ref{Chapter1} 

\lhead{\emph{Foreword}} % This is for the header on each page - perhaps a shortened title

%----------------------------------------------------------------------------------------
While I strongly support the axiom that scientific research should eschew first person pronouns, I'd like to offer a foreword encapsulating my own experiences with computer science, a narrative that I believe is highly relevant in a document examining the importance of K-12 computer science education. \par  
Coding is my greatest passion and my greatest asset. I've coded at every step of my career, from lab equipment for my physics thesis, databases for software companies, pollution data for environmental non-profits, and the list goes on. While my classmates struggled to find employment upon graduating in 2011 (the end of the recession), I immediately landed a cushy software job complete with a beer fridge and catered lunches. But perhaps most importantly: I code in my free time for \emph{for pleasure} because coding is an incredibly versatile creative platform. \par
And yet, upon graduating from high school, I had no idea what computer science entailed or that software engineering might be a career I'd later consider pursuing. If I had, I would have majored, or at least minored, in CS. In addition, I've written at length about the reasons why, \textit{as a woman}, I almost didn't major in physics in college. I'll suffice it to say, closing the gender gap in hard sciences begins in elementary and secondary education. \par
Fortunately, I lucked out and graduated with a degree that mandated ample coding experience, and I consider myself highly privileged to have found a subject that affords such opportunity. I hope to use my experiences not as a bias propelling a blind campaign in search of the latest trend, but rather as the bedrock passion behind the push to ensure that education keeps pace with evolving tools and means of exploration. \par 
\emph{Onwards!} \par
\emph{Jenna deBoisblanc}
