% Appendix Template

\chapter{Appendix: Online Curriculum} % Main appendix title

\label{AppendixTools} % Change X to a consecutive letter; for referencing this appendix elsewhere, use \ref{AppendixX}

\lhead{Appendix: \emph{Online Curriculum}} % Change X to a consecutive letter; this is for the header on each page - perhaps a shortened title

The following table is compiled from EdSurge\cite{edsurgetab} and Code.Org\cite{codeapps}. For full descritions and side-by-side comparisons, visit \href{https://code.org/educate/3rdparty}{Code.org's 3rd Educator Party Resources}.
\footnotesize
\begin{longtable}{p{2cm}p{4cm}p{3cm}p{1.3cm}p{1.3cm}p{1.3cm}}	
\caption{K-12 Online Curriculum} \\ \hline
\rowcolor[HTML]{D4DCE7} \textbf{Org} & \textbf{Curriculum} & \textbf{PD} & \textbf{Cost} & \textbf{Self-Guided?} & \textbf{Student Tracking?}  \\ \hline

%-----TABLE CONTENTS---------------------------------------%
\hline
\multicolumn{6}{l}{\cellcolor[HTML]{EFEFEF}\textbf{LOWER SCHOOL}} \\   \hline \hline
\href{https://studio.code.org/}{Code Studio (Code.org)} & 4 courses blend online tutorials with “unplugged” activities & 1-day weekend workshops across the US, FREE & FREE & Y & Y \\ \hline
\href{https://www.gethopscotch.com/}{Hopscotch} & This free iPad app uses a visual programming language similar to Scratch to help kids learn the basics of programming logic, such as sequencing, loops, variables, functions and conditionals. & - & FREE & N & N \\ \hline
\href{https://www.kodable.com/}{Kodable} & Kodable is a freemium educational iPad game offering a kid-friendly introduction to programming concepts and problem solving. & - & FREE & Y & Y \\ \hline
\href{https://www.pltw.org/}{Project Lead The Way} & 6, 10-hour computer science modules & Face-to-face and online, \$700 for school-level lead teacher & \$750 /school & Y & Y \\ \hline
\href{http://www.scratchjr.org/index.html}{Scratch Jr.} & An iPad app that introduces coding cocepts through games and visual programming. Assessments and curriculum included. & - & FREE & Y & N \\ \hline
\href{http://scratched.gse.harvard.edu/resources/search/results/taxonomy%3A21%2C28}{ScratchEd} & A 6-unit intro to Scratch, a visual programming language for game creation, animations, and art. & In-person educator meet-ups and online MOOC, FREE & FREE & N & N \\ \hline
\href{http://snap.berkeley.edu/}{SNAP!} & SNAP!'s visual blocks (similar to Scratch) support higher level computer science concepts like recursion, procedures, and continuations, SNAP! can work with LEGO Mindstorms NXT. & - & FREE & N & N \\ \hline
\href{https://www.tynker.com/}{Tynker} & Inspired by Scratch, Tynker has a dashboard to allow teachers to create a more structured way of teaching code with visual blocks. Includes assessment, classroom management, lesson plans, and a built in tutor. & 2-day PD, \$2000/day + travel & FREE tutorials. \$399 /class, \$2,000 /school & Y & Y \\ \hline

\\ \hline \hline
\multicolumn{6}{l}{\cellcolor[HTML]{EFEFEF}\textbf{MIDDLE SCHOOL}} \\ \hline \hline

\href{http://www.bootstrapworld.org/}{Bootstrap} & Teach algebra through video-game programming, with a 20-hr module to go alongside or inside a math class & 3-day workshops for schools and districts. Fees range & FREE & N & N \\ \hline
\href{https://code.org/educate}{Code.org} & 20-lesson CS in Algebra and CS in Science modules for math and science classes & In-person and online workshops available to partner districts, FREE, 50\% of teacher stipends covered by Code.org & FREE & Y & Y \\ \hline
\href{https://codehs.com/}{CodeHS} & Year-long Introduction to Computer Science and AP Computer Science in Java.  & Online PD for teaching intro computer science, 30-40 hour course, \$1000/teacher & FREE Intro; Pro version is \$2000/ section /year & Y & Y \\ \hline
\href{http://globaloria.com/courses-services/}{Globaloria} & 6 game-design courses & 3-day, in-person training and ongoing online PD, fee included in student price & \$75 /student & Y & Y \\ \hline
\href{https://nclab.com/karel/}{NCLab} & Courses in Karel coding, Turtle coding, 3D modeling, and Python. & Online & \$899 /class & Y & Y \\ \hline
\href{http://app.pluralsight.com/training/player?author=joe-hummel&name=learning-programming-scratch-m0-installation-windows&mode=live&clip=0&course=learning-programming-scratch}{Pluaralsight} & Free video-led coding courses for kids on Scratch, HTML, App Inventor, Kodu and Hopscotch & Online & FREE & Y & N \\ \hline
\href{https://www.pltw.org/}{Project Lead The Way} & 6, 10-hour computer science modules & Face-to-face and online, \$700 for school-level lead teacher & \$750 /school & Y & Y \\ \hline
\href{http://scratched.gse.harvard.edu/resources/search/results/taxonomy%3A21%2C28}{ScratchEd} & A 6-unit intro to Scratch & In-person educator meet-ups and online MOOC, FREE & FREE & N & N \\ \hline
\href{http://snap.berkeley.edu/}{SNAP!} & SNAP!'s visual blocks (similar to Scratch) support higher level computer science concepts like recursion, procedures, and continuations, SNAP! can work with LEGO Mindstorms NXT. & See Scratch & FREE & N & N \\ \hline
\href{https://www.tynker.com/}{Tynker} & Inspired by Scratch, Tynker has a dashboard to allow teachers to create a more structured way of teaching code with visual blocks. Includes assessment, classroom management, lesson plans, and a built in tutor. & 2-day PD, \$2000/day + travel & Free tutorials. \$399 /class, \$2,000 /school & Y & Y \\ \hline

\\ \hline \hline
\multicolumn{6}{l}{\cellcolor[HTML]{EFEFEF}\textbf{UPPER SCHOOL}} \\ \hline \hline

\href{http://bjc.berkeley.edu/}{Beauty and Joy of Computing} & Year-long CS Principles course & In-person in NYC, Berkeley, CA and North Carolina, FREE, stipends in NYC, stipends + travel elsewhere paid as available & FREE & N & N \\ \hline
\href{http://www.bootstrapworld.org/}{Bootstrap} & Teach algebra through video-game programming, with a 20-hr module to go alongside or inside a math class & 3-day workshops for schools and districts. Fees range & FREE & N & N \\ \hline
\href{https://www.codeavengers.com/teacher/info}{Code Avengers} & In-browser exercises and courses in JavaScript, HTML5, CSS3 and Python. Introductory courses are free, with intermediate and advanced courses for $29-$39.  & Online & FREE intro; intermediate and advanced courses for \$29-\$39.  & Y & Y \\ \hline
\href{https://code.org/educate}{Code.org} & Exploring Computer Science intro course and Computer Science Principles, a pilot course with an AP exam coming in 2016 & In-person and online workshops available to partner districts, FREE, 50\% of teacher stipends covered by Code.org & FREE & N & N \\ \hline
\href{https://www.codecademy.com/learn}{Codecademy} & 6 online courses (Javascript, SQL, Python, Ruby on Rails, Git, HTML/CSS, and more) and 65 projects. FREE & Not provided & FREE & Y & Y \\ \hline
\href{https://codehs.com/}{CodeHS} & Year-long Introduction to Computer Science and AP Computer Science in Java.  & Online PD for teaching intro computer science, 30-40 hour course, \$1000/teacher & FREE Intro; Pro version is \$2000/section/year & Y & Y \\ \hline
\href{https://edhesive.com/}{Edhesive} & Year-long AP Computer Science course & Online PD, community and content/technical/program support available, \$2,200 per school & FREE & Y & Y \\ \hline
\href{http://globaloria.com/courses-services/}{Globaloria} & 6 game-design courses & 3-day, in-person training and ongoing online PD, fee included in student price & \$75/student & Y & Y \\ \hline
\href{http://hello.processing.org/}{Hello Processing} & Video tutorials and interactive coding environment for learning Processing. & Online & FREE & Y & N \\ \hline
\href{https://www.khanacademy.org/computing/computer-science}{Khan Academy} & Online curriculum that teaches JavaScript programming, HTML/CSS, and SQL, in an interactive online environment, plus courses on Algorithms and Cryptography.  & Online & FREE & Y & Y \\ \hline
\href{https://ram8647.appspot.com/mobileCSP/}{Mobile CSP} & Year-long AP Computer Science Principles course based on App Inventor, a mobile programming language for Android devices. Materials available online & Online, regional in-person offered in CT, MA, NH and CA (others may be available), FREE, stipends available & FREE & Y & N \\ \hline
\href{https://nclab.com/karel/}{NCLab} & Courses in Karel coding, Turtle coding, 3D modeling, and Python. & Online & \$899 /class & Y & Y \\ \hline
\href{https://www.pltw.org/faq/computer-science}{Project Lead The Way} & Several courses: Intro CS, AP, Cybersecurity, and CS Applications & 5 or 10-day in-person training, \$1200 or \$2400, depending on course & \$2000 /school & Y & Y \\ \hline
\end{longtable}	

---

